\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2023-04-23 doc/USER_MANUAL/03_mesh_generation.tex

 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{setup/constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest this value as it represents a fair trade-off between element deformations to adapt to
topographic variations and numerical accuracy.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}.
Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
For higher numbers of doubling layers, additional entries must be added.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
The layer count starts from the bottom layer.
More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).\newline

>>>>

 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{setup/constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest this value as it represents a fair trade-off between element deformations to adapt to
topographic variations and numerical accuracy.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}.
Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
For higher numbers of doubling layers, additional entries must be added.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
The layer count starts from the bottom layer.
More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save \href{www.simulia.com}{Abaqus FEA} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save \href{www.opendx.org}{OpenDX} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).\newline

### 2023-04-23 doc/USER_MANUAL/04_creating_databases.tex

\vspace{1cm}
\noindent {\bf $\bullet$ Mesh setup}
\begin{description}
\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
elements. If you use the internal mesher, the only option is 8-node
bricks (27-node elements are not supported). \texttt{CUBIT} does not
support HEX27 elements either (it can generate them, but they are
flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
properly taken into account, you can use Gmsh \url{http://geuz.org/gmsh/}
\item [{\texttt{MODEL}}] Must be set to one of the following:

@@

save ParaView \urlwithparentheses{www.paraview.org} mesh files for

>>>>

\vspace{1cm}
\noindent {\bf $\bullet$ Mesh setup}
\begin{description}
\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
elements. If you use the internal mesher, the only option is 8-node
bricks (27-node elements are not supported). \texttt{CUBIT} does not
support HEX27 elements either (it can generate them, but they are
flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
properly taken into account, you can use Gmsh \url{http://gmsh.info}
\item [{\texttt{MODEL}}] Must be set to one of the following:

@@

save \href{www.paraview.org}{ParaView} mesh files for

### 2023-04-23 doc/USER_MANUAL/08_doing_tomography.tex

\red{\Large ANNOUNCEMENT (June 2017): See also the next chapter (Chapter~\ref{cha:FWI}) about how to perform full waveform inversion (FWI) or source inversions.}\newline

UNDER CONSTRUCTION (July 2015). New content will be added soon.

@@

One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
is to use them within a tomographic inversion, i.e., perform imaging. In other words, use
recorded seismograms, make measurements with synthetic seismograms,
and use the misfit between the two to iteratively improve the model
described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$ as a function of space.

As explained for instance in \cite{MoChKoWa15},
full waveform inversion means that one considers the observed seismograms (possibly filtered) as the
basic observables that one wants to fit. One thus searches for the model that minimizes the mean squared difference between observed
and synthetic seismograms. In other words, the goal is to find a structural model that can explain a larger portion
of seismological records, and not simply the phase of a few seismic arrivals.

@@

A direct method to compute the gradient is to take the derivative of (\ref{misfit}) with respect to model parameters:
\begin{equation}
\frac{\partial\chi(\boldm)}{\partial\boldm} = -\sum_{s=1}^N\sum_{r=1}^M \int_0^T
\frac{\partial \boldu(\boldx_r, \boldx_s; t)}{\partial\boldm}\cdot\left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right] \, dt \, .
\label{calc_gradient}
\end{equation}
This equation can be reformulated as the matrix-vector product:
\begin{equation}
\boldg = -\mathbf{J}^* \cdot \delta \mathbf{d} \, ,
\label{Jacobian}
\end{equation}
where $\mathbf{J}^*$ is the adjoint of the Jacobian matrix of the forward problem that contains the Fr\'echet derivatives of the
data with respect to model parameters, and $\delta \mathbf{d}$ is the vector that contains the data residuals.
The determination of $\mathbf{J}$ would require computing the Fr\'echet derivatives for each time step in the time
window considered and for all the source-station pairs, which is completely prohibitive on current supercomputers
(let us note that this situation may change one day).
However, it is possible to obtain this gradient without computing the Jacobian matrix explicitly.
The approach to determine the gradient without computing the Fr\'echet derivatives was introduced in nonlinear optimization
by \cite{Cha74} working with J. L. Lions, and later applied to seismic exploration problems by
\cite{BaChLa77}, \cite{BaChHeLa82}, \cite{Lai83} and \cite{Tar84}. The idea is to resort to the adjoint state,
which corresponds to the wavefield emitted and back-propagated from the receivers \cite[e.g.,][]{TrTaLi05,TrKoLi08,Plessix_2006_RAS,FiBuIg06}.

@@

As can be seen from these expressions, the principle of the adjoint-state method is to correlate two wavefields:
the direct (i.e. forward) field that propagates from the source to the receivers, and the adjoint field that propagates from all the
receivers backward in time. The same approach can be followed for any type of seismic observable (phase, amplitude,
envelope, time series...), provided the appropriate adjoint source is used \cite[]{TrTaLi05,TrKoLi08}.
For example, for the cross-correlated traveltime of a seismic
phase, the adjoint source is defined as the velocity of that synthetic phase weighted by the travel-time residual.

Computing the gradient based on the adjoint-state method requires performing two simulations per source (forward and adjoint
fields) regardless of the type of observable. However, to define the adjoint field one
must know the adjoint source, and that source is computed from the results of the forward simulation. One must therefore perform
the forward simulation before the adjoint simulation.
A straightforward solution for time-domain methods would be to store the whole forward field to disk at
each time step during the forward run and then read it back during the adjoint simulation to calculate the interaction of these two fields.
In 2-D this is feasible but in the 3-D case for very short seismic periods and without lossy compression, downsampling,
or a large amount of disk or memory checkpointing \cite[e.g.,][]{FiKeIg09,RuHaPuGu14,CyShWi15}
the amount of disk storage required would currently be too large.

However let us note again that this situation will change in the future.
In the mean time, a standard possible solution is to perform three simulations per source \cite[]{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
i.e., perform the forward calculation twice, once to compute the adjoint sources and once again
at the same time as the adjoint simulation to correlate the two fields and sum their interaction on the fly over all the time steps.
Doing so for an elastic Earth, one only needs a small amount of disk storage to store the last time
step of the forward run, which is then used as an initial condition to redo the forward run backwards,
as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.

@@

\begin{description}
\item [Smoothing:] To smooth a kernel, gradient or model file, use
\begin{verbatim}
mpirun -np NPROC bin/xsmooth_sem SIGMA_H SIGMA_V KERNEL_NAME INPUT_DIR OUPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{SIGMA\_H} horizontal smoothing radius,
\texttt{SIGMA\_V} vertical smoothing radius,
\texttt{KERNEL\_NAME} kernel name, e.g. \texttt{alpha\_kernel},
\texttt{INPUT\_DIR} directory from which kernels are read,
\texttt{OUTPUT\_DIR} directory to which smoothed kernels are written.

Smooths kernels by convolution with a Gaussian. Writes the resulting
smoothed kernels to \texttt{OUTPUT\_DIR}.

@@

\item [Clipping:] Values in a kernel, gradient or model binary file can be clipped with a minimum/maximum threshold value using
\begin{verbatim}
mpirun -np NPROC bin/xclip_sem MIN_VAL MAX_VAL KERNEL_NAMES INPUT_FILE OUTPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{MIN\_VAL} threshold below which array values are clipped,
\texttt{MAX\_VAL} threshold above which array values are clipped,
\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
\texttt{INPUT\_DIR} directory from which arrays are read,
\texttt{OUTPUT\_DIR} directory to which clipped array are written.

For each name in \texttt{KERNEL\_NAMES}, reads kernels from \texttt{INPUT\_DIR}, applies
thresholds, and writes the resulting clipped kernels to \texttt{OUTPUT\_DIR}.

@@

\item [Combining:] To combine kernel values from different kernel directories, you can use
\begin{verbatim}
mpirun -np NPROC bin/xcombine_sem KERNEL_NAMES INPUT_FILE OUTPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
\texttt{INPUT\_FILE} text file containing list of kernel directories,
\texttt{OUTPUT\_PATH} directory to which summed kernels are written.

For each name in \texttt{KERNEL\_NAMES}, sums kernels from directories specified in
\texttt{INPUT\_FILE}. Writes the resulting sums to \texttt{OUTPUT\_DIR}.

\texttt{INPUT\_FILE} is a text file containing a list of absolute or relative paths to
kernel directories, one directory per line.

@@

\begin{description}
\item [Isotropic model update:] The program \texttt{xadd\_model\_iso} can be used to update {\it isotropic} model files with
(smoothed \& summed) event kernels:
\begin{verbatim}
mpirun -np 4 bin/xadd_model_iso step_factor [INPUT-KERNELS-DIR/] [OUTPUT-MODEL-DIR/]
\end{verbatim}
with command line arguments:
\texttt{step\_factor} the step length to scale the gradient, e.g. $0.03$ for a $\pm 3$\% update,
\texttt{INPUT-KERNELS-DIR/} (optional) directory which holds summed kernels (e.g. \texttt{proc***alpha\_kernel.bin},.., by default
directory \texttt{INPUT\_GRADIENT/} is used),
\texttt{OUTPUT-MODEL-DIR/} (optional) directory which will hold new model files (e.g. \texttt{proc***vp\_new.bin},.., by default directory
\texttt{OUTPUT\_MODEL/} is used).

The gradients are given for isotropic parameters \texttt{(alpha,beta,rho)} or \texttt{(bulk\_c,beta,rho)}.

The algorithm uses a steepest descent method with a step length
determined by the given maximum update percentage.

@@

There are many other versions of gradient-based inversion
algorithms that could alternatively be used (see e.g. \cite{ViOp09,MoChKoWa15} for a list).
The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D\_Cartesian as well
as several additional components which are also stored on the CIG
svn server, described next.

The directory containing external utilities for tomographic inversion using
SPECFEM3D Cartesian (or other packages that evaluate misfit functions
and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
\begin{verbatim}
flexwin/     -- FLEXWIN algorithm for automated picking of time windows
measure_adj/ -- reads FLEXWIN output file and makes measurements,
                with the option for computing adjoint sources
iterate_adj/ -- various tools for iterative inversion
                (requires pre-computed "event kernels")
\end{verbatim}
This directory also contains
a brief \verb+README+ file indicating the role of the three subdirectories,
\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
The components for making the model update are there; however, there
are no explicit rules for computing the model update, just as with
any optimization problem. There are options for computing a conjugate
gradient step, as well as a source subspace projection step.

>>>>

\red{See also the next chapter (Chapter~\ref{cha:FWI}) about how to perform full waveform inversion (FWI) or source inversions.}\newline

STILL UNDER CONSTRUCTION ... New content will be added soon.

@@

One of the fundamental reasons for computing sensitivity kernels (Section~\ref{sec:Adjoint-simulation-finite})
is to use them within a tomographic inversion, i.e., perform imaging. In other words, use
recorded seismograms, make measurements with synthetic seismograms,
and use the misfit between the two to iteratively improve the model
described by (at least) $V_{{\rm p}}$, $V_{{\rm s}}$, and $\rho$ as a function of space.\newline

As explained for instance in \cite{MoChKoWa15},
full waveform inversion means that one considers the observed seismograms (possibly filtered) as the
basic observables that one wants to fit. One thus searches for the model that minimizes the mean squared difference between observed
and synthetic seismograms. In other words, the goal is to find a structural model that can explain a larger portion
of seismological records, and not simply the phase of a few seismic arrivals.\newline

@@

A direct method to compute the gradient is to take the derivative of (\ref{misfit}) with respect to model parameters:
\begin{equation}
\frac{\partial\chi(\boldm)}{\partial\boldm} = -\sum_{s=1}^N\sum_{r=1}^M \int_0^T
\frac{\partial \boldu(\boldx_r, \boldx_s; t)}{\partial\boldm}\cdot\left[\boldu(\boldx_r, \boldx_s; t) - \boldd(\boldx_r, \boldx_s; t)\right] \, dt \, .
\label{calc_gradient}
\end{equation}
This equation can be reformulated as the matrix-vector product:
\begin{equation}
\boldg = -\mathbf{J}^* \cdot \delta \mathbf{d} \, ,
\label{Jacobian}
\end{equation}
where $\mathbf{J}^*$ is the adjoint of the Jacobian matrix of the forward problem that contains the Fr\'echet derivatives of the
data with respect to model parameters, and $\delta \mathbf{d}$ is the vector that contains the data residuals.
The determination of $\mathbf{J}$ would require computing the Fr\'echet derivatives for each time step in the time
window considered and for all the source-station pairs, which is completely prohibitive on current supercomputers
(let us note that this situation may change one day).
However, it is possible to obtain this gradient without computing the Jacobian matrix explicitly.
The approach to determine the gradient without computing the Fr\'echet derivatives was introduced in nonlinear optimization
by \cite{Cha74} working with J. L. Lions, and later applied to seismic exploration problems by
\cite{BaChLa77}, \cite{BaChHeLa82}, \cite{Lai83} and \cite{Tar84}. The idea is to resort to the adjoint state,
which corresponds to the wavefield emitted and back-propagated from the receivers \cite[e.g.,][]{TrTaLi05,TrKoLi08,Plessix_2006_RAS,FiBuIg06}.\newline

@@

As can be seen from these expressions, the principle of the adjoint-state method is to correlate two wavefields:
the direct (i.e. forward) field that propagates from the source to the receivers, and the adjoint field that propagates from all the
receivers backward in time. The same approach can be followed for any type of seismic observable (phase, amplitude,
envelope, time series...), provided the appropriate adjoint source is used \cite[]{TrTaLi05,TrKoLi08}.
For example, for the cross-correlated traveltime of a seismic
phase, the adjoint source is defined as the velocity of that synthetic phase weighted by the travel-time residual.\newline

Computing the gradient based on the adjoint-state method requires performing two simulations per source (forward and adjoint
fields) regardless of the type of observable. However, to define the adjoint field one
must know the adjoint source, and that source is computed from the results of the forward simulation. One must therefore perform
the forward simulation before the adjoint simulation.
A straightforward solution for time-domain methods would be to store the whole forward field to disk at
each time step during the forward run and then read it back during the adjoint simulation to calculate the interaction of these two fields.
In 2-D this is feasible but in the 3-D case for very short seismic periods and without lossy compression, downsampling,
or a large amount of disk or memory checkpointing \cite[e.g.,][]{FiKeIg09,RuHaPuGu14,CyShWi15}
the amount of disk storage required would currently be too large.\newline

However let us note again that this situation will change in the future.
In the mean time, a standard possible solution is to perform three simulations per source \cite[]{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
i.e., perform the forward calculation twice, once to compute the adjoint sources and once again
at the same time as the adjoint simulation to correlate the two fields and sum their interaction on the fly over all the time steps.
Doing so for an elastic Earth, one only needs a small amount of disk storage to store the last time
step of the forward run, which is then used as an initial condition to redo the forward run backwards,
as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.\newline

@@

\begin{description}
\item [Smoothing:] To smooth a kernel, gradient or model file, use
\begin{verbatim}
mpirun -np NPROC bin/xsmooth_sem SIGMA_H SIGMA_V KERNEL_NAME INPUT_DIR OUPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{SIGMA\_H} horizontal smoothing radius,
\texttt{SIGMA\_V} vertical smoothing radius,
\texttt{KERNEL\_NAME} kernel name, e.g. \texttt{alpha\_kernel},
\texttt{INPUT\_DIR} directory from which kernels are read,
\texttt{OUTPUT\_DIR} directory to which smoothed kernels are written.\newline

This smooths kernels by convolution with a Gaussian, and writes the resulting
smoothed kernels to \texttt{OUTPUT\_DIR}.\newline

@@

\item [Clipping:] Values in a kernel, gradient or model binary file can be clipped with a minimum/maximum threshold value using
\begin{verbatim}
mpirun -np NPROC bin/xclip_sem MIN_VAL MAX_VAL KERNEL_NAMES INPUT_FILE OUTPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{MIN\_VAL} threshold below which array values are clipped,
\texttt{MAX\_VAL} threshold above which array values are clipped,
\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
\texttt{INPUT\_DIR} directory from which arrays are read,
\texttt{OUTPUT\_DIR} directory to which clipped array are written.\newline

For each name in \texttt{KERNEL\_NAMES}, reads kernels from \texttt{INPUT\_DIR}, applies
thresholds, and writes the resulting clipped kernels to \texttt{OUTPUT\_DIR}.\newline

@@

\item [Combining:] To combine kernel values from different kernel directories, you can use
\begin{verbatim}
mpirun -np NPROC bin/xcombine_sem KERNEL_NAMES INPUT_FILE OUTPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
\texttt{INPUT\_FILE} text file containing list of kernel directories,
\texttt{OUTPUT\_PATH} directory to which summed kernels are written.\newline

For each name in \texttt{KERNEL\_NAMES}, sums kernels from directories specified in
\texttt{INPUT\_FILE}. Writes the resulting sums to \texttt{OUTPUT\_DIR}.\newline

\texttt{INPUT\_FILE} is a text file containing a list of absolute or relative paths to
kernel directories, one directory per line.\newline

@@

\begin{description}
\item [Isotropic model update:] The program \texttt{xadd\_model\_iso} can be used to update {\it isotropic} model files with
(smoothed \& summed) event kernels:
\begin{verbatim}
mpirun -np 4 bin/xadd_model_iso step_factor [INPUT-KERNELS-DIR/] [OUTPUT-MODEL-DIR/]
\end{verbatim}
with command line arguments:
\texttt{step\_factor} the step length to scale the gradient, e.g. $0.03$ for a $\pm 3$\% update,
\texttt{INPUT-KERNELS-DIR/} (optional) directory which holds summed kernels (e.g. \texttt{proc***alpha\_kernel.bin},.., by default
directory \texttt{INPUT\_GRADIENT/} is used),
\texttt{OUTPUT-MODEL-DIR/} (optional) directory which will hold new model files (e.g. \texttt{proc***vp\_new.bin},.., by default directory
\texttt{OUTPUT\_MODEL/} is used).\newline

The gradients are given for isotropic parameters \texttt{(alpha,beta,rho)} or \texttt{(bulk\_c,beta,rho)}.\newline

The algorithm uses a steepest descent method with a step length
determined by the given maximum update percentage.\newline

@@

There are many other versions of gradient-based inversion
algorithms that could alternatively be used (see e.g. \cite{ViOp09,MoChKoWa15} for a list).
The tomographic inversion of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D\_Cartesian as well
as several additional components which are also stored on the CIG
svn server, described next.\newline

The directory containing external utilities for tomographic inversion using
SPECFEM3D Cartesian (or other packages that evaluate misfit functions
and gradients) is in directory \texttt{utils/ADJOINT\_TOMOGRAPHY\_TOOLS/}:
\begin{verbatim}
flexwin/     -- FLEXWIN algorithm for automated picking of time windows
measure_adj/ -- reads FLEXWIN output file and makes measurements,
                with the option for computing adjoint sources
iterate_adj/ -- various tools for iterative inversion
                (requires pre-computed "event kernels")
\end{verbatim}
This directory also contains
a brief \verb+README+ file indicating the role of the three subdirectories,
\verb+flexwin+ \citep{Maggi2009}, \verb+measure_adj+, and \verb+iterate_adj+.
The components for making the model update are there; however, there
are no explicit rules for computing the model update, just as with
any optimization problem. There are options for computing a conjugate
gradient step, as well as a source subspace projection step.\newline
